---
title: "Spatial Voting in US Presidential Election"
author: "Marie-Lou Sohnius"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2: 
    toc: false
  html_document: default
  pdf_document: default
documentclass: article
classoption: a4paper
subtitle: 'Quantitative Methods 2020: Final Data Essay'
abstract: |
  The text of your abstract.  150 -- 250 words.
keywords:
  - key
  - dictionary
  - word
linestretch: 2
geometry: "left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm"
---

```{r setup, include=FALSE}
# The first line sets an option for the final document that can be produced from
# the .Rmd file. Don't worry about it.
knitr::opts_chunk$set(echo = TRUE)

# The next bit is quite powerful and useful. 
# First you define which packages you need for your analysis and assign it to 
# the p_needed object. 
p_needed <-
  c("foreign", "viridis", "MASS", "optimx", 
    "separationplot", "knitr",
    "stargazer")

# Now you check which packages are already installed on your computer.
# The function installed.packages() returns a vector with all the installed 
# packages.
packages <- rownames(installed.packages())
# Then you check which of the packages you need are not installed on your 
# computer yet. Essentially you compare the vector p_needed with the vector
# packages. The result of this comparison is assigned to p_to_install.
p_to_install <- p_needed[!(p_needed %in% packages)]
# If at least one element is in p_to_install you then install those missing
# packages.
if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
# Now that all packages are installed on the computer, you can load them for
# this project. Additionally the expression returns whether the packages were
# successfully loaded.
sapply(p_needed, require, character.only = TRUE)

# This is an option for stargazer tables
# It automatically adapts the output to html or latex,
# depending on whether we want a html or pdf file
stargazer_opt <- ifelse(knitr::is_latex_output(), "latex", "html")

rm(list=ls())

set.seed(21122020)
```

```{r,  include=FALSE}
data <- read.dta("data/anes08_data_essay.dta")

vars <- c(1,3,8:15)

df <- data[complete.cases(data[ ,vars]),]


independents_hi <-
  df[which(df$party_id == "Independent" & df$pol_knowledge == 1), ]

mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```


```{r density, echo = FALSE, dpi=300, fig.width=9, fig.height=7, out.width='75%', fig.align='center'}

a <- density(df$lib_con_self[which(df$party_id == "Democrat")])
b <- density(df$lib_con_self[which(df$party_id == "Independent")])
c <- density(df$lib_con_self[which(df$party_id == "Republican")])

plot(a$x,
     a$y,
     xlab = "Liberal-Conservative Location of Voter",
     ylab = "Voter Density",
     ylim = c(min(c$y), max(c$y)),
     xlim = c(1, 7),
     col = "black",
     type = "l",
     lty = 1,
     lwd = 3,
     bty = "n",
     yaxt = "n"
     )

lines(b$x, b$y, col = "black", lty = 2, lwd = 3)

lines(c$x, c$y, col = "black", lty = 3, lwd = 3)

legend(
  "topright",
  legend = c("Democrat", "Independent", "Republican"),
  col = "black",
  lty = 1:3,
  cex = 0.8
)

```

```{r prop_voting, echo = FALSE, dpi=300, fig.width=9, fig.height=7, out.width='75%', fig.align='center'}

props_dem <- as.data.frame(prop.table(table(df$lib_con_self[which(df$party_id=="Democrat")], df$obama[which(df$party_id=="Democrat")])))
props_dem <- props_dem[which(props_dem$Var2==1),]

props_ind <- as.data.frame(prop.table(table(df$lib_con_self[which(df$party_id=="Independent")], df$obama[which(df$party_id=="Independent")])))
props_ind <- props_ind[which(props_ind$Var2==1),]

props_rep <- as.data.frame(prop.table(table(df$lib_con_self[which(df$party_id=="Republican")], df$obama[which(df$party_id=="Republican")])))
props_rep <- props_rep[which(props_rep$Var2==1),]


plot(
  c(1,7),
  ylim = c(0, 0.5),
  xlim = c(1, 7),
  type = "n",
  ylab = "Proportion Voting for Obama",
  xlab = "Liberal-Conservative Location of Voter",
  bty = "n",
  las = 1
)


lines(props_dem$Var1,
     props_dem$Freq, col = "black", lty = 1, lwd = 3)

lines(props_ind$Var1,
     props_ind$Freq, col = "black", lty = 2, lwd = 3)

lines(props_rep$Var1,
     props_rep$Freq, col = "black", lty = 3, lwd = 3)

legend(
  "topright",
  legend = c("Democrat", "Independent", "Republican"),
  col = "black",
  lty = 1:3,
  cex = 0.8
)


```



```{r,  include=FALSE}
m1 <-
  glm(
    obama ~ lib_con_self * party_id + race + male + age + relig_denomination + above_median_inc + college_degree,
    data = df,
    family = binomial(link = logit)
  )
summary(m1)


# 1. get the coefficients
beta_hat <- coef(m1)

# 2. Get the variance-covariance matrix
V_hat <- vcov(m1)

# 3. Set up a multivariate normal distribution N(beta_hat, V_hat)
# 4. Draw from the distribution nsim times
nsim <- 10000
S <- mvrnorm(n = nsim, 
             mu = beta_hat, 
             Sigma = V_hat)
head(S)

# we simulate over a sequence of x-values
lr_seq <- seq(min(df$lib_con_self, na.rm=T), max(df$lib_con_self, na.rm=T), length.out = 100)

# lr_seq <- seq(1,7,length.out = 7)
# What about the other covariates? Following average case practice, we take the 
# mode for dummy variables and rounding to the nearest whole number for categorical variables:



# Democrats 
scenario_dem <- cbind(1, # Intercept
                  lr_seq, # LR Scale
                  0, # Party ID = Independent
                  0, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  0, # LR Scale*Independent
                  0  # LR Scale*Republican
                  )

Xbeta_dem <- S %*% t(scenario_dem)

p_sim_dem <- (exp(Xbeta_dem))/ (1 + exp(Xbeta_dem))

p_mean_dem <- apply(p_sim_dem, 2, mean)
p_qu_dem <- t(apply(p_sim_dem, 2, quantile, prob = c(0.025, 0.975)))

# Independents 
scenario_ind <- cbind(1, # Intercept
                  lr_seq, # LR Scale
                  1, # Party ID = Independent
                  0, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  lr_seq*1, # LR Scale*Independent
                  0  # LR Scale*Republican
                  )

Xbeta_ind <- S %*% t(scenario_ind)

p_sim_ind <- (exp(Xbeta_ind))/ (1 + exp(Xbeta_ind))

p_mean_ind <- apply(p_sim_ind, 2, mean)
p_qu_ind <- t(apply(p_sim_ind, 2, quantile, prob = c(0.025, 0.975)))

# Republicans 
scenario_rep <- cbind(1, # Intercept
                  lr_seq, # LR Scale
                  0, # Party ID = Independent
                  1, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  0, # LR Scale*Independent
                  lr_seq*1  # LR Scale*Republican
                  )

Xbeta_rep <- S %*% t(scenario_rep)

p_sim_rep <- (exp(Xbeta_rep))/ (1 + exp(Xbeta_rep))

p_mean_rep <- apply(p_sim_rep, 2, mean)
p_qu_rep <- t(apply(p_sim_rep, 2, quantile, prob = c(0.025, 0.975)))


```


```{r ev_partyid, echo = FALSE, dpi=300, fig.width=7, fig.height=7, out.width='75%', fig.align='center'}

plot(
  x = lr_seq,
  y = p_mean_dem,
  ylim = c(0, 1),
  type = "n",
  main = "Party Identification and Spatial Voting",
  ylab = "Probability of Voting for Obama",
  xlab = "Liberal-Conservative Location of Voter",
  bty = "n",
  las = 1
)

# Democrats

# plot uncertainty with a polygon
polygon(x = c(rev(lr_seq), lr_seq),
        y = c(rev(p_qu_dem[,2]), p_qu_dem[,1]),
        col = adjustcolor("black", alpha = 0.1),
        border = NA)

# and a line
lines(x = lr_seq,
      y = p_mean_dem,
      lwd = 3,
      lty = 1)

# Independents

# plot uncertainty with a polygon
polygon(x = c(rev(lr_seq), lr_seq),
        y = c(rev(p_qu_ind[,2]), p_qu_ind[,1]),
        col = adjustcolor("black", alpha = 0.1),
        border = NA)

# and a line
lines(x = lr_seq,
      y = p_mean_ind,
      lwd = 3,
      lty = 2)


# Republicans

# plot uncertainty with a polygon
polygon(x = c(rev(lr_seq), lr_seq),
        y = c(rev(p_qu_rep[,2]), p_qu_rep[,1]),
        col = adjustcolor("black", alpha = 0.1),
        border = NA)

# and a line
lines(x = lr_seq,
      y = p_mean_rep,
      lwd = 3,
      lty = 3)


legend(
  "bottomleft",
  legend = c("Democrat", "Independent", "Republican"),
  col = c("black", "black", "black"),
  lty = 1:3,
  cex = 0.8,
  pt.lwd = 2
)

abline(h = 0.5, col = adjustcolor("red", alpha = 0.5), lty = 1, lwd = 3) 

```


```{r,  include=FALSE}

obama_position <- 2.83
mccain_position <- 5.28

# Democrats

scenario_dem_1 <- cbind(1, # Intercept
                  obama_position, # LR Scale
                  0, # Party ID = Independent
                  0, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  0, # LR Scale*Independent
                  0  # LR Scale*Republican
                  )


Xbeta_dem_1 <- S %*% t(scenario_dem_1)

p_sim_dem_1 <- (exp(Xbeta_dem_1))/ (1 + exp(Xbeta_dem_1))

p_mean_dem_1 <- apply(p_sim_dem_1, 2, mean)
p_qu_dem_1 <- t(apply(p_sim_dem_1, 2, quantile, prob = c(0.025, 0.975)))



scenario_dem_2 <- cbind(1, # Intercept
                  mccain_position, # LR Scale
                  0, # Party ID = Independent
                  0, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  0, # LR Scale*Independent
                  0  # LR Scale*Republican
                  )

Xbeta_dem_2 <- S %*% t(scenario_dem_2)

p_sim_dem_2 <- (exp(Xbeta_dem_2))/ (1 + exp(Xbeta_dem_2))

p_mean_dem_2 <- apply(p_sim_dem_2, 2, mean)
p_qu_dem_2 <- t(apply(p_sim_dem_2, 2, quantile, prob = c(0.025, 0.975)))


fd_dem <- p_sim_dem_1 - p_sim_dem_2
fd_mean_dem <- apply(fd_dem, 2, mean)
fd_qu_dem <- t(apply(fd_dem, 2, quantile, prob = c(0.025, 0.975)))

# Independents

scenario_ind_1 <- cbind(1, # Intercept
                  obama_position, # LR Scale
                  1, # Party ID = Independent
                  0, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  obama_position*1, # LR Scale*Independent
                  0  # LR Scale*Republican
                  )

Xbeta_ind_1 <- S %*% t(scenario_ind_1)

p_sim_ind_1 <- (exp(Xbeta_ind_1))/ (1 + exp(Xbeta_ind_1))

p_mean_ind_1 <- apply(p_sim_ind_1, 2, mean)
p_qu_ind_1 <- t(apply(p_sim_ind_1, 2, quantile, prob = c(0.025, 0.975)))




scenario_ind_2 <- cbind(1, # Intercept
                  mccain_position, # LR Scale
                  1, # Party ID = Independent
                  0, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  mccain_position*1, # LR Scale*Independent
                  0  # LR Scale*Republican
                  )

Xbeta_ind_2 <- S %*% t(scenario_ind_2)

p_sim_ind_2 <- (exp(Xbeta_ind_2))/ (1 + exp(Xbeta_ind_2))

p_mean_ind_2 <- apply(p_sim_ind_2, 2, mean)
p_qu_ind_2 <- t(apply(p_sim_ind_2, 2, quantile, prob = c(0.025, 0.975)))

fd_ind <- p_sim_ind_1 - p_sim_ind_2
fd_mean_ind <- apply(fd_ind, 2, mean)
fd_qu_ind <- t(apply(fd_ind, 2, quantile, prob = c(0.025, 0.975)))


# Republicans

scenario_rep_1 <- cbind(1, # Intercept
                  obama_position, # LR Scale
                  0, # Party ID = Independent
                  1, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  0, # LR Scale*Independent
                  obama_position*1  # LR Scale*Republican
                  )


Xbeta_rep_1 <- S %*% t(scenario_rep_1)

p_sim_rep_1 <- (exp(Xbeta_rep_1))/ (1 + exp(Xbeta_rep_1))

p_mean_rep_1 <- apply(p_sim_rep_1, 2, mean)
p_qu_rep_1 <- t(apply(p_sim_rep_1, 2, quantile, prob = c(0.025, 0.975)))


scenario_rep_2 <- cbind(1, # Intercept
                  mccain_position, # LR Scale
                  0, # Party ID = Independent
                  1, # Party ID = Republican
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(df$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0, # College Degree
                  0, # LR Scale*Independent
                  mccain_position*1  # LR Scale*Republican
                  )

Xbeta_rep_2 <- S %*% t(scenario_rep_2)

p_sim_rep_2 <- (exp(Xbeta_rep_2))/ (1 + exp(Xbeta_rep_2))

p_mean_rep_2 <- apply(p_sim_rep_2, 2, mean)
p_qu_rep_2 <- t(apply(p_sim_rep_2, 2, quantile, prob = c(0.025, 0.975)))

fd_rep <- p_sim_rep_1 - p_sim_rep_2
fd_mean_rep <- apply(fd_rep, 2, mean)
fd_qu_rep <- t(apply(fd_rep, 2, quantile, prob = c(0.025, 0.975)))

fds <- rbind(fd_mean_dem,fd_mean_ind, fd_mean_rep)
cis <- rbind(fd_qu_dem,fd_qu_ind, fd_qu_rep)


```

```{r, echo = FALSE, dpi=300, fig.width=7, fig.height=3.5, out.width='100%', fig.align='center', fig.cap='FDs'}


# Plot
plot(
  cis,
  xlim = c(0.5,3.5),
  ylim = c(0, 0.5),
  type = "n",
  xlab = "",
  xaxt = "n",
  ylab = "First Difference",
  bty = "n",
  main = "FDs"
)

axis(1, at=1:3, labels=c("Democrat", "Independent", "Republican"))

abline(h = 0, col = "#eea1a1", lty = 1, lwd = 3)

# Now we add the confidence intervals from our 100 regressions on different samples.
col <- c("black", "#989898", "#cccccc")

for (i in 1:3) {
  segments(i,cis[i, 1], i, cis[i, 2], col = col[i], lwd = 4)
  points(i,
         fds[i],
         pch = 19,
         cex = 1,
         lwd = 4,
         col = col[i])
}



```

```{r,  include=FALSE}


m2 <-
  glm(obama ~ lib_con_self * party_id,
      data = df,
      family = binomial(link = logit))
summary(m2)


```

```{r, echo = FALSE, dpi=300, fig.width=7, fig.height=3.5, out.width='100%', fig.align='center', fig.cap='Model fit'}

par(2,1)
separationplot(pred = m2$fitted, 
               actual = df$obama,
               line = TRUE,
               heading = "Separation Plot", 
               show.expected = T,
               col0 = "#cccccc",
               col1 = "#989898",
               lwd2 = 2,
               height = 2,
               newplot = F)

separationplot(pred = m1$fitted, 
               actual = df$obama,
               line = TRUE,
               heading = "Separation Plot", 
               show.expected = T,
               col0 = "#cccccc",
               col1 = "#989898",
               lwd2 = 2,
               height = 2,
               newplot = F)
```


```{r,  include=FALSE}


m3 <-
  glm(
    obama ~ lib_con_self + race + male + age + relig_denomination + above_median_inc + college_degree,
    data = independents_hi,
    family = binomial(link = logit)
  )
summary(m3)



# 1. get the coefficients
beta_hat <- coef(m3)

# 2. Get the variance-covariance matrix
V_hat <- vcov(m3)

# 3. Set up a multivariate normal distribution N(beta_hat, V_hat)
# 4. Draw from the distribution nsim times
nsim <- 10000
S <- mvrnorm(n = nsim, 
             mu = beta_hat, 
             Sigma = V_hat)
head(S)

# we simulate over a sequence of x-values
lr_seq2 <- seq(1, 7, by = 0.001)

# lr_seq <- seq(1,7,length.out = 7)
# What about the other covariates? Following average case practice, we take the 
# mode for dummy variables and rounding to the nearest whole number for categorical variables:



# Scenario 
scenario_ind_hi <- cbind(1, # Intercept
                  lr_seq2, # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  1, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi <- S %*% t(scenario_ind_hi)

p_sim_ind_hi <- (exp(Xbeta_ind_hi))/ (1 + exp(Xbeta_ind_hi))

p_mean_ind_hi <- apply(p_sim_ind_hi, 2, mean)
p_qu_ind_hi <- t(apply(p_sim_ind_hi, 2, quantile, prob = c(0.025, 0.975)))

props_ind_hi <- as.data.frame(cbind(p_mean_ind_hi,lr_seq2))
props_ind_hi$dist <- abs(props_ind_hi$p_mean_ind_hi-0.5)
min_pos <- which.min(props_ind_hi$dist)
props_ind_hi[min_pos,2]


# Position=4 
scenario_ind_hi2 <- cbind(1, # Intercept
                  4, # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  1, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi2 <- S %*% t(scenario_ind_hi2)

p_sim_ind_hi2 <- (exp(Xbeta_ind_hi2))/ (1 + exp(Xbeta_ind_hi2))

p_mean_ind_hi2 <- apply(p_sim_ind_hi2, 2, mean)
p_qu_ind_hi2 <- t(apply(p_sim_ind_hi2, 2, quantile, prob = c(0.025, 0.975)))



# Prop = 0.5
scenario_ind_hi3 <- cbind(1, # Intercept
                  props_ind_hi[min_pos,2], # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  1, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi3 <- S %*% t(scenario_ind_hi3)

p_sim_ind_hi3 <- (exp(Xbeta_ind_hi3))/ (1 + exp(Xbeta_ind_hi3))

p_mean_ind_hi3 <- apply(p_sim_ind_hi3, 2, mean)
p_qu_ind_hi3 <- t(apply(p_sim_ind_hi3, 2, quantile, prob = c(0.025, 0.975)))


fd <- p_sim_ind_hi2 - p_sim_ind_hi3

# calculate means and quantiles
fd_mean <- apply(fd, 2, mean)
fd_qu <- t(apply(fd, 2, quantile, prob = c(0.025, 0.975)))



```

```{r ev_informed, echo = FALSE, dpi=300, fig.width=9, fig.height=7, out.width='75%', fig.align='center'}

plot(
  x = lr_seq2,
  y = p_mean_ind_hi,
  ylim = c(0, 1),
  type = "n",
  main = "Party Identification and Spatial Voting",
  ylab = "Probability of Voting for Obama",
  xlab = "Liberal-Conservative Location of Voter",
  bty = "n",
  las = 1
)

# Democrats

# plot uncertainty with a polygon
polygon(x = c(rev(lr_seq2), lr_seq2),
        y = c(rev(p_qu_ind_hi[,2]), p_qu_ind_hi[,1]),
        col = adjustcolor("black", alpha = 0.2),
        border = NA)

# and a line
lines(x = lr_seq2,
      y = p_mean_ind_hi,
      lwd = 3,
      lty = 1)
```
```{r}



# Scenario (Black)
scenario_ind_hi_r1 <- cbind(1, # Intercept
                  lr_seq2, # LR Scale
                  0, # Race = Black
                  1, # Race = Other/Mixed
                  1, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi_r1 <- S %*% t(scenario_ind_hi_r1)

p_sim_ind_hi_r1 <- (exp(Xbeta_ind_hi_r1))/ (1 + exp(Xbeta_ind_hi_r1))

p_mean_ind_hi_r1 <- apply(p_sim_ind_hi_r1, 2, mean)
p_qu_ind_hi_r1 <- t(apply(p_sim_ind_hi_r1, 2, quantile, prob = c(0.025, 0.975)))

props_ind_hi_r1 <- as.data.frame(cbind(p_mean_ind_hi_r1,lr_seq2))
props_ind_hi_r1$dist <- abs(props_ind_hi_r1$p_mean_ind_hi_r1-0.5)
min_pos <- which.min(props_ind_hi_r1$dist)
props_ind_hi_r1[min_pos,2]


# Scenario 2 (Female)
scenario_ind_hi_r2 <- cbind(1, # Intercept
                  lr_seq2, # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  0, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi_r2 <- S %*% t(scenario_ind_hi_r2)

p_sim_ind_hi_r2 <- (exp(Xbeta_ind_hi_r2))/ (1 + exp(Xbeta_ind_hi_r2))

p_mean_ind_hi_r2 <- apply(p_sim_ind_hi_r2, 2, mean)
p_qu_ind_hi_r2 <- t(apply(p_sim_ind_hi_r2, 2, quantile, prob = c(0.025, 0.975)))

props_ind_hi_r2 <- as.data.frame(cbind(p_mean_ind_hi_r2,lr_seq2))
props_ind_hi_r2$dist <- abs(props_ind_hi_r2$p_mean_ind_hi_r2-0.5)
min_pos <- which.min(props_ind_hi_r2$dist)
props_ind_hi_r2[min_pos,2]


# Scenario 3 (Young)
scenario_ind_hi_r3 <- cbind(1, # Intercept
                  lr_seq2, # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  1, # Gender = Male
                  quantile(independents_hi$age, na.rm=T, 0.25), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi_r3 <- S %*% t(scenario_ind_hi_r3)

p_sim_ind_hi_r3 <- (exp(Xbeta_ind_hi_r3))/ (1 + exp(Xbeta_ind_hi_r3))

p_mean_ind_hi_r3 <- apply(p_sim_ind_hi_r3, 2, mean)
p_qu_ind_hi_r3 <- t(apply(p_sim_ind_hi_r3, 2, quantile, prob = c(0.025, 0.975)))

props_ind_hi_r3 <- as.data.frame(cbind(p_mean_ind_hi_r3,lr_seq2))
props_ind_hi_r3$dist <- abs(props_ind_hi_r3$p_mean_ind_hi_r3-0.5)
min_pos <- which.min(props_ind_hi_r3$dist)
props_ind_hi_r3[min_pos,2]



# Scenario 4 (No Religion)
scenario_ind_hi_r4 <- cbind(1, # Intercept
                  lr_seq2, # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  1, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  0, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi_r4 <- S %*% t(scenario_ind_hi_r4)

p_sim_ind_hi_r4 <- (exp(Xbeta_ind_hi_r4))/ (1 + exp(Xbeta_ind_hi_r4))

p_mean_ind_hi_r4 <- apply(p_sim_ind_hi_r4, 2, mean)
p_qu_ind_hi_r4 <- t(apply(p_sim_ind_hi_r4, 2, quantile, prob = c(0.025, 0.975)))

props_ind_hi_r4 <- as.data.frame(cbind(p_mean_ind_hi_r4,lr_seq2))
props_ind_hi_r4$dist <- abs(props_ind_hi_r4$p_mean_ind_hi_r4-0.5)
min_pos <- which.min(props_ind_hi_r4$dist)
props_ind_hi_r4[min_pos,2]



# Scenario (Low Income)
scenario_ind_hi_r5 <- cbind(1, # Intercept
                  lr_seq2, # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  1, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  0, # Above Median Income
                  0 # College Degree
                  )

Xbeta_ind_hi_r5 <- S %*% t(scenario_ind_hi_r5)

p_sim_ind_hi_r5 <- (exp(Xbeta_ind_hi_r5))/ (1 + exp(Xbeta_ind_hi_r5))

p_mean_ind_hi_r5 <- apply(p_sim_ind_hi_r5, 2, mean)
p_qu_ind_hi_r5 <- t(apply(p_sim_ind_hi_r5, 2, quantile, prob = c(0.025, 0.975)))

props_ind_hi_r5 <- as.data.frame(cbind(p_mean_ind_hi_r5,lr_seq2))
props_ind_hi_r5$dist <- abs(props_ind_hi_r5$p_mean_ind_hi_r5-0.5)
min_pos <- which.min(props_ind_hi_r5$dist)
props_ind_hi_r5[min_pos,2]


# Scenario 6 (College Degree)
scenario_ind_hi_r6 <- cbind(1, # Intercept
                  lr_seq2, # LR Scale
                  0, # Race = Black
                  0, # Race = Other/Mixed
                  1, # Gender = Male
                  mean(independents_hi$age, na.rm=T), # Age
                  1, # Religion = Protestant
                  0, # Religion = Catholic
                  0, # Religion = Other
                  1, # Above Median Income
                  1 # College Degree
                  )

Xbeta_ind_hi_r6 <- S %*% t(scenario_ind_hi_r6)

p_sim_ind_hi_r6 <- (exp(Xbeta_ind_hi_r6))/ (1 + exp(Xbeta_ind_hi_r6))

p_mean_ind_hi_r6 <- apply(p_sim_ind_hi_r6, 2, mean)
p_qu_ind_hi_r6 <- t(apply(p_sim_ind_hi_r6, 2, quantile, prob = c(0.025, 0.975)))

props_ind_hi_r6 <- as.data.frame(cbind(p_mean_ind_hi_r6,lr_seq2))
props_ind_hi_r6$dist <- abs(props_ind_hi_r6$p_mean_ind_hi_r6-0.5)
min_pos <- which.min(props_ind_hi_r6$dist)
props_ind_hi_r6[min_pos,2]




```


```{r ev_informed_robust, echo = FALSE, dpi=300, fig.width=9, fig.height=7, out.width='75%', fig.align='center'}

plot(
  x = lr_seq2,
  y = p_mean_ind_hi,
  ylim = c(0, 1),
  type = "n",
  main = "Party Identification and Spatial Voting",
  ylab = "Probability of Voting for Obama",
  xlab = "Liberal-Conservative Location of Voter",
  bty = "n",
  las = 1
)


lines(x = lr_seq2,
      y = p_mean_ind_hi_r1,
      lwd = 3,
      lty = 1)

lines(x = lr_seq2,
      y = p_mean_ind_hi_r2,
      lwd = 3,
      lty = 2)

lines(x = lr_seq2,
      y = p_mean_ind_hi_r3,
      lwd = 3,
      lty = 3)

lines(x = lr_seq2,
      y = p_mean_ind_hi_r4,
      lwd = 3,
      lty = 4)

lines(x = lr_seq2,
      y = p_mean_ind_hi_r5,
      lwd = 3,
      lty = 5)

lines(x = lr_seq2,
      y = p_mean_ind_hi_r6,
      lwd = 3,
      lty = 6)
```


```{r, echo = FALSE, results='asis'}


stargazer(
  m2,
  m1,
  m3,
  # You can also add more than just one model
  type = stargazer_opt,
  # Important to produce a nice table in the output document
  covariate.labels = c(
    "Left-Right",
    "Independent",
    "Republican",
    "Black/African-American",
    "Other/Mixed",
    "Male",
    "Age",
    "Protestant",
    "Catholic",
    "Other",
    "Above Median Income",
    "College Degree",
    "Left-Right $\\times$ Independent",
    "Left-Right $\\times$ Republican"
  ),
  dep.var.labels = c("Vote for Obama"),
  column.labels=c("Model 1\\\ (Base)", "Model 1\\\ (with controls)", "Model 2"),
  header = F,
  label = "tab:Models",
  title = "Logit equations predicting the vote"
)
```

# Introduction

As you know by now, this is an R Markdown document. By this point in the semester, you should be familiar with how this works. We still want to give you some tips if you want to write your data essay in R Markdown.

We want you to submit your data essay in .pdf format. You can generate .pdf files from your .Rmd files. This is done by selecting `Knit` > `Knit to pdf`. If this does not work yet on yur computer, run the following code:


# Organizing the write-up with section titles

You can use `#` hashtags to organize you write-up with section-titles on different levels. This is how it works:

# Section

## Subsection

## Subsubsection

### Subsubsubsection

Markdown can do many more things: 

## Text formatting options

Markdown also provides an easy way to format the text in your document. Here are some basics:

  - *This text is italic*
  - **This text is bold**
  - `This is code`
  - [This is a link to the R Markdown homepage](https://rmarkdown.rstudio.com)

For more formatting options, we recommend the [R Markdown Cheat Sheet](https://rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf?_ga=2.195559076.94966807.1602002830-1180689241.1600075670).

# Chunk Options

As you know, you can write code in R Markdown. This means, that you can do the coding and the write up in one single document. This is a great feature because it ensures that it will be easy to reproduce your document right away. However, to get a good looking write up, you most likely (especially if it should like in a journal article) only want to show tables and figures but hide the code in your final document. This is not difficult to achieve in R Markdown. You only have to adjust chunk options.

Chunk output in your final document can be customized with knitr options, arguments set in the `{}` of a chunk header. Here are some useful arguments:

  - `include = FALSE` prevents code and results from appearing in the finished file. R Markdown still runs the code in the chunk, and the results can be used by other chunks. This is useful for running your data pre-processing and models.
  - `echo = FALSE` prevents code, but not the results from appearing in the finished file. This is a useful way to embed figures.
  - `message = FALSE` prevents messages that are generated by code from appearing in the finished file.
  - `warning = FALSE` prevents warnings that are generated by code from appearing in the finished.
  - `fig.cap = "..."` adds a caption to a figure produced by the chunk.

See the [R Markdown Reference Guide](https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf?_ga=2.169319849.94966807.1602002830-1180689241.1600075670) for a complete list of knitr chunk options.

# Example

Let's do an example: We want to simulate some data and create a plot. But the write up should only show the plot without any code.

```{r, include=FALSE}
# we simulate some data
# althoug we want to run this code, it should not appear in the write-up
# that's why we include 'include=FALSE' in the chunk options above

x <- runif(1000, 1, 4)
y <- runif(1000, 1, 4)

col <- round(x*y)
```

```{r single-plot, echo = FALSE, dpi=300, fig.width=7, fig.height=7, out.width='75%', fig.align='center', fig.cap='A colorful plot'}

# we set a couple of options:

# echo=FALSE makes sure that we show the output of the chunk (but not the code)
# dpi=300 makes sure that the resolution of the plot is high enough
# fig.width=7 defines the relative with of the plot
# fig.height=7 defines the relative height of the plot
# out.width='75%' defines the total width of the plot
# fig.align='center' defines the alignment of the plot
# fig.cap='A colorful plot' generates a caption for the plot

plot(x = x,
     y = y,
     xlim = c(1, 4),
     ylim = c(1, 4),
     xaxt = "n",
     yaxt = "n",
     pch = 19,
     bty = "n",
     xlab = "",
     ylab = "",
     col = viridis(max(col), 0.5)[col])
```

You can even cross-reference Figures in your text. For this to work you need to install the bookdown package and set the output to bookdown::pdf_document2. You can see how to do this in the yaml header of this document. Instead of knit to pdf_document you can then use knit to pdf_document2. 

You just have to set a chunk name and use it like: In Figure \@ref(fig:single-plot) you can see very colorful points. You can also include calculated values directly in your text. The minimum value on the x-axis is `r round(min(x), 2)` and the mean value on the x-axis is `r round(mean(x), 2)`.

You can also plot two (or more) plots next to each other. You can see these in Figure \@ref(fig:two-plots).^[This is a footnote.]

```{r two-plots, echo = FALSE, dpi=300, fig.width=7, fig.height=3.5, out.width='100%', fig.align='center', fig.cap='Two colorful plots'}

# we set a couple of options:

# echo=FALSE makes sure that we show the output of the chunk (but not the code)
# dpi=300 makes sure that the resolution of the plot is high enough
# fig.width=7 defines the relative with of the plot
# fig.height=3.5 defines the relative height of the plot
# out.width='100%' defines the total width of the plot
# fig.align='center' defines the alignment of the plot
# fig.cap='Two colorful plots' generates a caption for the plot

# generate a 1 by 2 plot window
par(mfrow = c(1, 2),
    # we reduce the blank space around the plot 
    # (where axis labels and titles are usually plotted)
    mar = c(0, 0, 0, 0) + 0.1) 

plot(x = x,
     y = y,
     xlim = c(1, 4),
     ylim = c(1, 4),
     xaxt = "n",
     yaxt = "n",
     pch = 19,
     bty = "n",
     xlab = "",
     ylab = "",
     col = viridis(max(col), 0.5)[col])

plot(x = x,
     y = y,
     xlim = c(1, 4),
     ylim = c(1, 4),
     xaxt = "n",
     yaxt = "n",
     pch = 22,
     cex = 2,
     bty = "n",
     xlab = "",
     ylab = "",
     col = viridis(round(max(x)), 0.5)[round(x)])
```

\pagebreak

You also might to show a table (e.g. regression table) and reference Table \@ref(tab:regression) in your text.

```{r, echo = FALSE, results='asis'}

# we set three options here:

# echo=FALSE makes sure that we show the output of the chunk (but not the code)
# message=FALSE hides all messages created in the code chunk
# warning=FALSE hides all warnings created in the code chunk
# results='asis' makes sure that the table gets printed nicely

# note that for this to work, you must define "stargazer_opt" in the setup chunk
# as we did in this file (see line 45).

linear_model <- lm(y ~ x)

stargazer(linear_model, # You can also add more than just one model
          type = stargazer_opt, # Important to produce a nice table in the output document
          covariate.labels = c("X"), # Make sure to have the labels in the right order
          dep.var.labels = c("Y"),
          header = F, # To surpress the stargazer comment
          label = "tab:regression", # If you produce pdf output you can even cross-reference your table in your text.
          title = "Regression Table")
```

Feel free to get in touch if you have a question regarding your write-up.

**Good Luck  with the Data Essay!**
